C/O and Snowline Locations in Protoplanetary Disks: The Effect of Radial Drift and Viscous Gas Accretion

### 2015-10-05 tex/draft/ms_final_revised.tex
update

\begin{abstract}
%The C/O ratio is an important signature of giant planet atmospheric composition and disk chemistry. 
The C/O ratio is a defining feature of both gas giant atmospheric and protoplanetary disk chemistry. In disks, the C/O ratio is regulated by the presence of snowlines of major volatiles at different distances from the central star.
We explore the effect of radial drift of solids and viscous gas accretion onto the central star on the snowline locations of the main C and O carriers in a protoplanetary disk, H$_2$O, CO$_2$ and CO, and their consequences for the C/O ratio in gas and dust throughout the disk. We determine the snowline locations for a range of fixed initial particle sizes, in both an active and a passive disk. We find that grains with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for a passive disk and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an active disk desorb %instantaneously and [I think instantaneously is too strong]
at a %fixed 
size-dependent location in the disk, which is independent %regardless 
of the particle's initial position. The snowline radius decreases for larger particles, up to sizes of $\sim$7~m. %location moves inwards as the particle size increases, and the innermost snowlines are set by grains with initial size $s\sim7$ m. 
Compared to a static disk, we find that radial drift and gas accretion \textbf{in a steady-state accretionally heated disk} move the H$_2$O snowline inwards by up to \textbf{40 \%}, the CO$_2$ snowlines by up to 60 \%, and the CO snowline by up to 50 \%. We thus determine an inner limit on the snowline locations when radial drift and gas accretion are accounted for. %Our results have direct implications on the C/O ratio in giant planet atmospheres that have formed in situ. [I don't think we discuss this enough to warrant it as a conclusion
\end{abstract}

@@

The structures of protoplanetary disks are complex, and affected by a multitude of chemical and dynamical processes (see review by \citealt{henning13}). From the chemistry perspective, volatile compounds are particularly important. Their snowline locations determine their relative abundance in gaseous and solid form in the disk,. Based on protostellar and comet abundances, some of the most important volatile molecules are H$_2$O, CO$_2$, CO, N$_2$. Recent observations of protoplanetary disks have provided valuable information about the abundances and snowline locations of some of these compounds. For example, the CO snowline has been detected in the disk around TW Hya \citep{qi13}. Observations of TW Hya have also revealed a H$_2$O snowline \citep{zhang13}, and more such snowline detections are expected in future ALMA cycles. These observations are currently lacking an interpretive framework that takes into account all important dynamical and chemical processes. Furthermore, such a framework is crucial to connect observed snowline locations to planet formation.

@@

An important consequence of snowline formations in disks is that disks are expected to present different carbon-to-oxygen (C/O) ratios in the gas and in icy dust mantles at different disk radii. This effect was quantified by  \citet{oberg11}, who considered the fact that the main carries of carbon and oxygen, i.e. H$_2$O, CO$_2$ and CO, have different condensation temperatures. This changes the relative abundance of C and O in gaseous and solid form as a function of the snowline location of the volatiles mentioned above. \citet{oberg11} calculated analytically the C/O ratio in gas in dust as a function of semimajor axis for passive protoplanetary disks and found a gas C/O ratio of order unity between the CO$_2$ and CO snowlines, where oxygen gas is highly depleted. This effect was used to explain claims of detections of superstellar C/O ratios in exoplanet atmospheres (e.g., WASP-12b, \citealt{madhu11}). 
%Superstellar C/O ratios have also been detected in the atmospheres of gas giants such as WASP-12b \citep{madhu11}.

@@

\citet{oberg11} assumed a static disk with no chemical evolution. In reality, dynamical and chemical processes affect the snowline locations and the resulting C/O ratio. Several works have addressed some of these effects. \citet{madhu14} use a steady-state active disk model that includes planetary migration and use the C/O ratio to constrain migration mechanisms. \citet{alidib14} calculate the C/O ratio throughout the disk by incorporating the evolution of solids, i.e. radial drift, sublimation and grain coagulation, as well as the diffusion of volatile vapors. \citet{alidib14} use the 1+1D $\alpha$-disk model of \citet{hughes10}, in which the gas drifts outwards in the disk midplane, and thus small particles that are well-coupled to the gas will also advect outward. Their model assumes a cyclical conversion between H$_2$O or CO dust and vapor: large enough particles that are decoupled from the gas drift inwards and start desorbing. Once their sublimation is complete, back-diffusion moves the H$_2$O or CO vapor ouwards to their respective snowlines, where they instantly condense into mm-sized particles that diffuse outwards with the gas while coagulating into larger particles. Once the grains become large enough to decouple from the gas and drift inwards, the cycle restarts. This ```conveyor belt''' model is also used by \citet{ciesla06} for the evolution of H$_2$O in a viscous disk. This approach leads \citet{alidib14} to find that the gaseous C/O ratio increases with time inside the H$_2$O snowline, approaching unity at 2 AU after $\sim$$10^4-10^5$ years.  \citet{thiabaud15} consider additional carbon carrier volatile species in their chemical network, such as CH$_4$, and find that the gas C/O ratio may be enriched by up to four times the Solar value in the outer parts of the disk where CH$_4$ and CO are the only gaseous carriers of C and O.  They also include nitrogen carriers such as N$_2$ or NH$_3$, and perform similar calculations for nitrogen.

%Most studies of the C/O ratio in disks are particularly focused on the inner disk, in order to explain observations of super-stellar C/O ratios in the atmospheres of close-in exoplanets \citep{madhu11}. In addition, the simultaneous inclusion of various dynamical and/or chemical processes makes it difficult to quantify the separate effect of each of these processes on snowline locations and the C/O ratio. %[The preceding paragraph needs to be made clearer. Currently it is not clear what effects these studies have had on our understanding of C/O ratio sin disks and why your study is needed. I.e. why are the processes you focus on the most important ones? Introduce separately papers that deal with C/O and water snowline locations.]
Each of these studies have considered a specific combination of dynamical and chemical effects. One combination that has not yet been considered is the combination of radial drift and viscous gas accretion. Studying these two dynamical processes makes it possible to quantify their separate effect on snowline locations and the C/O ratio at various disk radii.

@@

This paper is organized as follows. In Section \ref{sec:model}, we present our disk, radial drift and desorption models, as well as the timescales relevant to the coupled drift-desorption process. We calculate the H$_2$O, CO$_2$ and CO snowline locations as a function of particle size for a passive and an active disk in Section \ref{sec:snowlines}, and the resulting C/O ratio throughout the disk in Section \ref{sec:COratio}. In Section \ref{sec:discussion}, we discuss the generality of our results, as well as additional effects on the snowline locations. Finally, we summarize our findings in Section \ref{sec:summary}.

@@

We present our protoplanetary disk model for both a passive and an active disk in section \ref{sec:disk}. In section \ref{sec:drift}, we describe our analytic model for the radial drift of solids. We summarize our ice desorption model in section \ref{sec:desorption}. Finally, we discuss the relevant timescales for dynamical effects that affect snowline locations %in the desorption process 
in section \ref{sec:timescales}.

@@

To understand the separate effects of radial drift, radial movement of gas throughout the disk due to gas accretion, and accretion heating, we use four separate disk models:  \textit{passive disk}, which is only irradiated by the central star and does not experience gas accretion or accretional heating; \textit{active disk with passive temperature profile} (hereafter \textit{active disk}), in which the gas is accreting onto the central star causing the gas surface density to decrease with time, but which does not experience accretion heating; \textit{active disk at steady-state with accretion heating} (hereafter \textit{active disk accretionally heated}), for which the mass flux $\dot{M}$ is constant in time and independent of semimajor axis and the temperature profile is calculated using both accretional heating and stellar irradiation; and \textit{static disk}, which has a passive temperature profile and does not take into account gas accretion onto the central star or radial drift.

@@

\textbf{Passive disk.} We adopt a minimum mass solar nebula (MMSN) disk model for a passive disk similar to the prescription of \citet{chiang10}. The gas surface density and midplane temperature are
\begin{subeqnarray}
\label{eq:disk}
\Sigma&=&2000\, (r/\text{AU})^{-1}\,\, \text{g cm}^{-2} \slabel{eq:disksigma}\\
T &=& 120\, (r/\text{AU})^{-3/7} \,\,\text{K}, \slabel{eq:diskT}
\end{subeqnarray}
where $r$ is the semimajor axis. Our surface density profile is flatter than the $\Sigma \propto r^{-3/2}$ used by \citet{chiang10}. Our choice is inspired by observations of protoplanetary disks at radii larger than $\sim$20 AU (e.g., \citealt{andrews10}), which suggest that typical disks may have surface density profiles with $\Sigma \propto r^{-1}$.

@@

\textbf{Active disk.} We model the active disk as a thin disk with an $\alpha$-viscosity prescription \citep{shakura73}:
\begin{equation}
\label{eq:nu}
\nu=\alpha c H.
\end{equation}
Here $\nu$ is the kinematic viscosity, $\alpha < 1$ is a dimensionless coefficient and we choose $\alpha=0.01$, and $c$, $H$ are the isothermal sound speed and disk scale height, respectively:
\begin{subeqnarray}
\label{eq:cdHd}
c &=& \sqrt{\frac{k_{\rm B} T}{\mu m_{\rm p}}} \slabel{eq:cd} \\
H&=& \frac{c}{\Omega_{\rm k}} \slabel{eq:Hd},
\end{subeqnarray}
where $k_{\rm B}$ is the Boltzmann constant, $\mu$ is the mean molecular weight of the gas, $m_{\rm p}$ is the proton mass, and $\Omega_{\rm k} \equiv \sqrt{G M_*/r^3}$ is the Keplerian angular velocity,  with $G$ the gravitational constant and $M_*$ the stellar mass. We choose $M_*=M_{\odot}$ and $\mu=2.35$, corresponding to the Solar composition of hydrogen and helium. The temperature profile for the %an 
active disk is assumed to be the same as for the passive disk and given by Equation (\ref{eq:diskT}). From Equations (\ref{eq:nu}) and (\ref{eq:cdHd}), the viscosity can thus be expressed as a power-law in radius, $\nu \propto r^{\gamma}$, with $\gamma=15/14 \approx 1$ for our choice of parameters. Following \citet{hartmann98}, we define $R \equiv r/r_{\rm c}$ and $\nu_{\rm c} \equiv \nu(r_{\rm c})$, where $r_{\rm c}$ is a characteristic disk radius. We choose $r_{\rm c}=100$ AU. The gas surface density is given by the self-similar solution

@@

\textbf{Active disk accretionally heated.} %A steady-state solution to the evolution of an actively accreting disk gives a gas surface density $\Sigma$ and mass flux $\dot{M}$ that are constant in time and independent of semimajor axis. Since 
Calculating the midplane temperature self-consistently for an active disk is non-trivial. We thus use instead the Shakura-Sunyaev thin disk steady-state solution to derive the midplane temperature profile, $T_{\rm act}$. The equations governing the evolution of the steady-state disk are listed in \App{app:steadystate}. We assume an interstellar opacity for the dust grains given by \citet{bell94}, but reduced by a factor of 100. This reduction is due to the fact that disk opacities are lower than the interstellar one. While this scaling is consistent with more detailed models of grain opacities in disks (e.g., \citealt{mordasini14}), realistic disk opacities are much less sensitive to changes in temperature than the interstellar opacity if substantial grain growth has occurred. However, the disk temperature does not vary significantly across the small region of the disk where accretion heating is important ($r \lesssim 1$ AU). Moreover, using an analytic opacity formula is more convenient since it results in a constant gas surface density in the inner disk region (see below). Our opacity law is thus

\begin{equation}
\label{eq:opacity}
\kappa=\kappa_0 T_{\rm act}^2,
\end{equation}
where $\kappa_0=2 \times 10^{-6}$. By solving the Equation set (\ref{eq:diskeq}) we find
\begin{equation}
\label{eq:Tdact}
T_{\rm act}=\frac{1}{4 r} \Big(\frac{3 G \kappa_0\dot{M}^2 M_* \mu m_{\rm p} \Omega_{\rm k}}{\pi^2 \alpha k_{\rm B} \sigma}\Big)^{1/3}.
\end{equation}
Since both accretion heating and stellar irradiation contribute to the thermal evolution of the disk, we compute the midplane temperature for our steady-state active disk as
\begin{equation}
\label{eq:activeT}
T^4 = T_{\rm act}^4 + T_{\rm pas}^4,
\end{equation}
where to avoid notation confusion $T_{\rm pas}=T$ from Equation (\ref{eq:diskT}), the temperature profile for a passive disk. We can then easily determine  $c$ and $H$ from Equation (\ref{eq:cdHd}), as well as the viscosity $\nu$ from Equation (\ref{eq:nu}) for a given $\alpha$. For consistency, we choose  $\alpha=0.01$ as in the previous case. Finally, we determine $\Sigma$ from Equation (\ref{eq:Mdot}), where we choose $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$ based on disk observations (e.g., \citealt{andrews10}). In the inner portion of our disk ($r  \lesssim 1$ AU for our fiducial model with $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$), our choice of opacity (Equation \ref{eq:opacity}) implies that the disk has a constant surface density with radius (see Equations \ref{eq:diskeq}). %We note that $\Sigma$ is no longer constant when stellar irradiation is taken into account in the active disk steady-state solution.

@@

For a passive disk, the radial drift velocity can be approximated as
\begin{equation}
\label{eq:rdotpas}
\dot{r} \approx -2 \eta \Omega_{\rm k} r \Big(\frac{\tau_{\rm s}}{1+\tau_{\rm s}^2}\Big),
\end{equation}
where
\begin{equation}
\label{eq:eta}
\eta \equiv - \frac{\partial P/\partial \ln r}{2 \rho v_{\rm k}^2} \approx \frac{c^2}{2 v_k^2}
\end{equation}
and $P = \rho c^2$ is the disk midplane pressure.

For an active disk, the radial drift velocity has an additional term due to the radial movement of the gas \citep{birnstiel12}, i.e.
\begin{equation}
\label{eq:rdotact}
\dot{r} \approx -2 \eta \Omega_{\rm k} r \Big(\frac{\tau_{\rm s}}{1+\tau_{\rm s}^2}\Big) + \frac{\dot{r}_{\rm gas}}{1+\tau_{\rm s}^2},
\end{equation}
where $\dot{r}_{\rm gas}$ is the radial gas accretion velocity and can be expressed as (e.g., \citealt{fkr02})
\begin{equation}
\label{eq:vgas}
\dot{r}_{\rm gas} = - \frac{3}{\Sigma \sqrt{r}} \frac{\partial}{\partial r}(\nu \Sigma \sqrt{r}) 
\end{equation}
with $\Sigma$ from Equation (\ref{eq:Sigmaact}). For the active disk steady-state solution (see Section \ref{sec:disk}), $\dot{r}_{\rm gas}$ can be expressed more simply using the definition of the mass flux, $\dot{M}=-2 \pi r \dot{r}_{\rm gas} \Sigma$, with $\dot{M}$ fixed and $\Sigma$ obtained from Equation (\ref{eq:Mdot}). \textbf{For our choice of parameters for both active disks, we have found that the radial flow of gas is always directed inward, in contrast with the model of \citet{alidib14} which assumes that the gas drifts outwards (see Section \ref{sec:intro}). We note that variations in our fiducial disk model parameters (e.g., $T$, $\Sigma$, $\dot{M}$) may cause the gas to flow outwards in the outer parts of the disk, specifically at the CO$_2$ and CO snowline locations. Since drifting particles larger than a few cm are only modestly affected by gas accretion, an outward gas flow would move the CO$_2$ and CO locations further away from the star only for the smallest particles in our model, which are well-coupled to the gas.}

@@

\textit{Radial drift timescale.} To order of magnitude, the radial drift timescale can be estimated as 
\begin{equation}
\label{eq:tdrift}
t_{\rm drift} \sim \Big|\frac{r}{\dot{r}}\Big|,
\end{equation}
where $\dot{r}$ is the radial drift velocity given by Equation (\ref{eq:rdotpas}) for a passive disk and by Equation (\ref{eq:rdotact}) for an active disk.

\textit{Gas accretion timescale.} The timescale for gas accretion onto the central star for an active disk is (e.g., \citealt{armitage10})
\begin{equation}
\label{eq:tgas}
t_{\rm gas, acc} \sim \frac{r^2}{\nu} \sim \frac{1}{2 \alpha \eta \Omega_{\rm k}},
\end{equation}
with the latter expression derived from Equations (\ref{eq:nu}) and (\ref{eq:eta}).

@@

For simplicity, we calculate the radial drift timescale, $t_{\rm drift}$, for a passive disk in this section, but most of our conclusions hold true for an active disk as well. Figure \ref{fig:timescales} shows $t_{\rm des}$, $t_{\rm drift}$ and $t_{\rm gas, acc}$ as a function of particle size at three different locations in the disk, corresponding to the H$_2$O, CO$_2$ and CO snowlines in the static disk. As expected, micron-sized particles desorb on very short timescales of $\sim 1-1000$ years in the close vicinity of their respective snowlines, since the desorption rate depends exponentially on temperature and hence on disk location (see Equation \ref{eq:Rdes}).  On the other hand, their radial drift timescale exceeds the typical disk lifetime of a few Myr by several orders of magnitude due to their strong coupling with the gas. Thus for small particles in a passive disk, the snowline locations and the C/O ratio are the same as for a static disk (see Figure 1 from \citealt{oberg11}). This is not true for an active disk, however, where gas accretion causes even micron-sized particles to drift significantly before desorbing, as we show in section \ref{sec:snowlines}. At the other extreme, kilometer-sized particles are unaffected by gas drag and have long desorption timescales ($\gg$1 Myr ), and the snowline locations and C/O ratio remain unchanged in this case as well. This is true for both passive and active disks, since large planetesimals are decoupled from the gas and hence unaffected by gas accretion onto the host star.

>>>>

\begin{abstract}
%The C/O ratio is an important signature of giant planet atmospheric composition and disk chemistry. 
The C/O ratio is a defining feature of both gas giant atmospheric and protoplanetary disk chemistry. In disks, the C/O ratio is regulated by the presence of snowlines of major volatiles at different distances from the central star.
We explore the effect of radial drift of solids and viscous gas accretion onto the central star on the snowline locations of the main C and O carriers in a protoplanetary disk, H$_2$O, CO$_2$ and CO, and their consequences for the C/O ratio in gas and dust throughout the disk. We determine the snowline locations for a range of fixed initial particle sizes, in both \textbf{evolving (active)} and \textbf{irradiated (passive) disks}. \textbf{For our fiducial disk model,} we find that grains with sizes $\sim$$0.5$ cm $\lesssim s \lesssim$ 7 m for an irradiated disk and $\sim$$0.001$ cm $\lesssim s \lesssim$ 7 m for an evolving disk desorb %instantaneously and [I think instantaneously is too strong]
at a %fixed 
size-dependent location in the disk, which is independent %regardless 
of the particle's initial position. The snowline radius decreases for larger particles, up to sizes of $\sim$7~m. %location moves inwards as the particle size increases, and the innermost snowlines are set by grains with initial size $s\sim7$ m. 
Compared to a static disk, we find that radial drift and gas accretion \textbf{in a viscous disk} move the H$_2$O snowline inwards by up to \textbf{40 \%}, the CO$_2$ snowlines by up to 60 \%, and the CO snowline by up to 50 \%. We thus determine an inner limit on the snowline locations when radial drift and gas accretion are accounted for. %Our results have direct implications on the C/O ratio in giant planet atmospheres that have formed in situ. [I don't think we discuss this enough to warrant it as a conclusion
\end{abstract}

@@

The structures of protoplanetary disks are complex, and affected by a multitude of chemical and dynamical processes (see review by \citealt{henning13}). From the chemistry perspective, volatile compounds are particularly important. Their snowline locations determine their relative abundance in gaseous and solid form in the disk,. Based on protostellar and comet abundances, some of the most important volatile molecules are H$_2$O, CO$_2$, CO, N$_2$. Recent observations of protoplanetary disks have provided valuable information about the abundances and snowline locations of some of these compounds. For example, the CO snowline has been detected in the disk around TW Hya \citep{qi13}, \textbf{as well as in the disk around HD 163296 \citep{mathews13} using line emissions from DCO$^+$}. Observations of TW Hya have also revealed a H$_2$O snowline \citep{zhang13}, and more such snowline detections are expected in future ALMA cycles. These observations are currently lacking an interpretive framework that takes into account all important dynamical and chemical processes. Furthermore, such a framework is crucial to connect observed snowline locations to planet formation.

@@

An important consequence of snowline formations in disks is that disks are expected to present different carbon-to-oxygen (C/O) ratios in the gas and in icy dust mantles at different disk radii. This effect was quantified by  \citet{oberg11}, who considered the fact that the main carries of carbon and oxygen, i.e. H$_2$O, CO$_2$ and CO, have different condensation temperatures. This changes the relative abundance of C and O in gaseous and solid form as a function of the snowline location of the volatiles mentioned above. \citet{oberg11} calculated analytically the C/O ratio in gas in dust as a function of semimajor axis for passive protoplanetary disks and found a gas C/O ratio of order unity between the CO$_2$ and CO snowlines, where oxygen gas is highly depleted. This effect was used to explain claims of detections of superstellar C/O ratios in exoplanet atmospheres (e.g., WASP-12b, \citealt{madhu11}), \textbf{which however have been unambiguously refuted (\citealt{stevenson14}, \citealt{kreidberg15})}.
%Superstellar C/O ratios have also been detected in the atmospheres of gas giants such as WASP-12b \citep{madhu11}.

@@

\citet{oberg11} assumed a static disk with no chemical evolution. In reality, dynamical and chemical processes affect the snowline locations and the resulting C/O ratio. Several works have addressed some of these effects. \citet{madhu14} use a steady-state active disk model that includes planetary migration and use the C/O ratio to constrain migration mechanisms. \citet{alidib14} calculate the C/O ratio throughout the disk by incorporating the evolution of solids, i.e. radial drift, sublimation and grain coagulation, as well as the diffusion of volatile vapors. \citet{alidib14} use the 1+1D $\alpha$-disk model of \citet{hughes10}, in which the gas drifts outwards in the disk midplane, and thus small particles that are well-coupled to the gas will also advect outward. Their model assumes a cyclical conversion between H$_2$O or CO dust and vapor: large enough particles that are decoupled from the gas drift inwards and start desorbing. Once their sublimation is complete, back-diffusion moves the H$_2$O or CO vapor ouwards to their respective snowlines, where they instantly condense into mm-sized particles that diffuse outwards with the gas while coagulating into larger particles. Once the grains become large enough to decouple from the gas and drift inwards, the cycle restarts. This ```conveyor belt''' model is \textbf{based on the pioneering work by \citet{cuzzi04}}, and \citet{ciesla06} for the evolution of H$_2$O in a viscous disk. This approach leads \citet{alidib14} to find that the gaseous C/O ratio increases with time inside the H$_2$O snowline, approaching unity at 2 AU after $\sim$$10^4-10^5$ years.  \citet{thiabaud15} consider additional carbon carrier volatile species in their chemical network, such as CH$_4$, and find that the gas C/O ratio may be enriched by up to four times the Solar value in the outer parts of the disk where CH$_4$ and CO are the only gaseous carriers of C and O.  They also include nitrogen carriers such as N$_2$ or NH$_3$, and perform similar calculations for nitrogen.

%Most studies of the C/O ratio in disks are particularly focused on the inner disk, in order to explain observations of super-stellar C/O ratios in the atmospheres of close-in exoplanets \citep{madhu11}. In addition, the simultaneous inclusion of various dynamical and/or chemical processes makes it difficult to quantify the separate effect of each of these processes on snowline locations and the C/O ratio. %[The preceding paragraph needs to be made clearer. Currently it is not clear what effects these studies have had on our understanding of C/O ratio sin disks and why your study is needed. I.e. why are the processes you focus on the most important ones? Introduce separately papers that deal with C/O and water snowline locations.]
Each of these studies have considered a specific combination of dynamical and chemical effects. One \textbf{scenario} that has not yet been considered is the combination of radial drift and viscous gas accretion \textbf{in isolation}. Studying these two dynamical processes makes it possible to quantify their separate effect on snowline locations and the C/O ratio at various disk radii.

@@

This paper is organized as follows. In Section \ref{sec:model}, we present our disk, radial drift and desorption models, as well as the timescales relevant to the coupled drift-desorption process. We calculate the H$_2$O, CO$_2$ and CO snowline locations as a function of particle size for \textbf{an irradiated and an evolving disk} in Section \ref{sec:snowlines}, and the resulting C/O ratio throughout the disk in Section \ref{sec:COratio}. In Section \ref{sec:discussion}, we discuss the generality of our results, as well as additional effects on the snowline locations. Finally, we summarize our findings in Section \ref{sec:summary}.

@@

We present our protoplanetary disk model for \textbf{an irradiated, an evolving, and a viscous disk} in section \ref{sec:disk}. In section \ref{sec:drift}, we describe our analytic model for the radial drift of solids. We summarize our ice desorption model in section \ref{sec:desorption}. Finally, we discuss the relevant timescales for dynamical effects that affect snowline locations %in the desorption process 
in section \ref{sec:timescales}.

@@

To understand the separate effects of radial drift, radial movement of gas throughout the disk due to gas accretion, and accretion heating, we use four separate disk models:  \textit{static disk}, which is solely irradiated by the host star and does not take into account gas accretion onto the star or radial drift; \textit{\textbf{irradiated} disk}, which has the same temperature profile as the static disk and does not experience gas accretion or accretional heating, but it takes into account radial drift of solids; \textit{\textbf{evolving} disk}, in which the gas is accreting onto the central star causing the gas surface density to decrease with time, but which does not experience accretion heating; and \textit{\textbf{viscous} disk}, for which the mass flux $\dot{M}$ is constant in time and independent of semimajor axis and the temperature profile is calculated using both accretional heating and stellar irradiation.

@@

\textbf{Irradiated disk.} We adopt a minimum mass solar nebula (MMSN) disk model for a irradiated disk similar to the prescription of \citet{chiang10}. The gas surface density and midplane temperature are
\begin{subeqnarray}
\label{eq:disk}
\Sigma&=&2000\, (r/\text{AU})^{-1}\,\, \text{g cm}^{-2} \slabel{eq:disksigma}\\
T &=& 120\, (r/\text{AU})^{-3/7} \,\,\text{K}, \slabel{eq:diskT}
\end{subeqnarray}
where $r$ is the semimajor axis. Our surface density profile is flatter than the $\Sigma \propto r^{-3/2}$ used by \citet{chiang10}. Our choice is inspired by observations of protoplanetary disks at radii larger than $\sim$20 AU (e.g., \citealt{andrews10}), which suggest that typical disks may have surface density profiles with $\Sigma \propto r^{-1}$.

@@

\textbf{Evolving disk.} We model the evolving disk as a thin disk with an $\alpha$-viscosity prescription \citep{shakura73}:
\begin{equation}
\label{eq:nu}
\nu=\alpha c H.
\end{equation}
Here $\nu$ is the kinematic viscosity, $\alpha < 1$ is a dimensionless coefficient and we choose $\alpha=0.01$, and $c$, $H$ are the isothermal sound speed and disk scale height, respectively:
\begin{subeqnarray}
\label{eq:cdHd}
c &=& \sqrt{\frac{k_{\rm B} T}{\mu m_{\rm p}}} \slabel{eq:cd} \\
H&=& \frac{c}{\Omega_{\rm k}} \slabel{eq:Hd},
\end{subeqnarray}
where $k_{\rm B}$ is the Boltzmann constant, $\mu$ is the mean molecular weight of the gas, $m_{\rm p}$ is the proton mass, and $\Omega_{\rm k} \equiv \sqrt{G M_*/r^3}$ is the Keplerian angular velocity,  with $G$ the gravitational constant and $M_*$ the stellar mass. We choose $M_*=M_{\odot}$ and $\mu=2.35$, corresponding to the Solar composition of hydrogen and helium. The temperature profile for the %an 
evolving disk is assumed to be the same as for the irradiated disk and given by Equation (\ref{eq:diskT}). From Equations (\ref{eq:nu}) and (\ref{eq:cdHd}), the viscosity can thus be expressed as a power-law in radius, $\nu \propto r^{\gamma}$, with $\gamma=15/14 \approx 1$ for our choice of parameters. Following \citet{hartmann98}, we define $R \equiv r/r_{\rm c}$ and $\nu_{\rm c} \equiv \nu(r_{\rm c})$, where $r_{\rm c}$ is a characteristic disk radius. We choose $r_{\rm c}=100$ AU. The gas surface density is given by the self-similar solution

@@

\textbf{Viscous disk.} %A steady-state solution to the evolution of an actively accreting disk gives a gas surface density $\Sigma$ and mass flux $\dot{M}$ that are constant in time and independent of semimajor axis. Since 
Calculating the midplane temperature self-consistently for an evolving disk is non-trivial. We thus use instead the Shakura-Sunyaev thin disk steady-state solution to derive the midplane temperature profile, $T_{\rm act}$. The equations governing the evolution of the steady-state disk are listed in \App{app:steadystate}. We assume an interstellar opacity for the dust grains given by \citet{bell94}, but reduced by a factor of 100. This reduction is due to the fact that disk opacities are lower than the interstellar one. While this scaling is consistent with more detailed models of grain opacities in disks (e.g., \citealt{mordasini14}), realistic disk opacities are much less sensitive to changes in temperature than the interstellar opacity if substantial grain growth has occurred. However, the disk temperature does not vary significantly across the small region of the disk where accretion heating is important ($r \lesssim 1$ AU). Moreover, using an analytic opacity formula is more convenient since it results in a constant gas surface density in the inner disk region (see below). Our opacity law is thus

\begin{equation}
\label{eq:opacity}
\kappa=\kappa_0 T_{\rm act}^2,
\end{equation}
where $\kappa_0=2 \times 10^{-6}$. By solving the Equation set (\ref{eq:diskeq}) we find
\begin{equation}
\label{eq:Tdact}
T_{\rm act}=\frac{1}{4 r} \Big(\frac{3 G \kappa_0\dot{M}^2 M_* \mu m_{\rm p} \Omega_{\rm k}}{\pi^2 \alpha k_{\rm B} \sigma}\Big)^{1/3}.
\end{equation}
Since both accretion heating and stellar irradiation contribute to the thermal evolution of the disk, we compute the midplane temperature for our viscous disk as
\begin{equation}
\label{eq:activeT}
T^4 = T_{\rm act}^4 + T_{\rm pas}^4,
\end{equation}
where to avoid notation confusion $T_{\rm pas}=T$ from Equation (\ref{eq:diskT}), the temperature profile for an irradiated disk. We can then easily determine  $c$ and $H$ from Equation (\ref{eq:cdHd}), as well as the viscosity $\nu$ from Equation (\ref{eq:nu}) for a given $\alpha$. For consistency, we choose  $\alpha=0.01$ as in the previous case. Finally, we determine $\Sigma$ from Equation (\ref{eq:Mdot}), where we choose $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$ based on disk observations (e.g., \citealt{andrews10}). In the inner portion of our disk ($r  \lesssim 1$ AU for our fiducial model with $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$), our choice of opacity (Equation \ref{eq:opacity}) implies that the disk has a constant surface density with radius (see Equations \ref{eq:diskeq}). %We note that $\Sigma$ is no longer constant when stellar irradiation is taken into account in the active disk steady-state solution.

@@

For an irradiated disk, the radial drift velocity can be approximated as
\begin{equation}
\label{eq:rdotpas}
\dot{r} \approx -2 \eta \Omega_{\rm k} r \Big(\frac{\tau_{\rm s}}{1+\tau_{\rm s}^2}\Big),
\end{equation}
where
\begin{equation}
\label{eq:eta}
\eta \equiv - \frac{\partial P/\partial \ln r}{2 \rho v_{\rm k}^2} \approx \frac{c^2}{2 v_k^2}
\end{equation}
and $P = \rho c^2$ is the disk midplane pressure.

For an evolving disk, the radial drift velocity has an additional term due to the radial movement of the gas \citep{birnstiel12}, i.e.
\begin{equation}
\label{eq:rdotact}
\dot{r} \approx -2 \eta \Omega_{\rm k} r \Big(\frac{\tau_{\rm s}}{1+\tau_{\rm s}^2}\Big) + \frac{\dot{r}_{\rm gas}}{1+\tau_{\rm s}^2},
\end{equation}
where $\dot{r}_{\rm gas}$ is the radial gas accretion velocity and can be expressed as (e.g., \citealt{fkr02})
\begin{equation}
\label{eq:vgas}
\dot{r}_{\rm gas} = - \frac{3}{\Sigma \sqrt{r}} \frac{\partial}{\partial r}(\nu \Sigma \sqrt{r}) 
\end{equation}
with $\Sigma$ from Equation (\ref{eq:Sigmaact}). For the viscous disk (see Section \ref{sec:disk}), $\dot{r}_{\rm gas}$ can be expressed more simply using the definition of the mass flux, $\dot{M}=-2 \pi r \dot{r}_{\rm gas} \Sigma$, with $\dot{M}$ fixed and $\Sigma$ obtained from Equation (\ref{eq:Mdot}). \textbf{For our choice of parameters for both the evolving and the viscous disks, we have found that the radial flow of gas is always directed inward, in contrast with the model of \citet{alidib14} which assumes that the gas drifts outwards (see Section \ref{sec:intro}). We note that variations in our fiducial disk model parameters (e.g., $T$, $\Sigma$, $\dot{M}$) may cause the gas to flow outwards in the outer parts of the disk, specifically at the CO$_2$ and CO snowline locations. Since drifting particles larger than a few cm are only modestly affected by gas accretion, an outward gas flow would move the CO$_2$ and CO locations further away from the star only for the smallest particles in our model, which are well-coupled to the gas.}

@@

\textit{Radial drift timescale.} To order of magnitude, the radial drift timescale can be estimated as 
\begin{equation}
\label{eq:tdrift}
t_{\rm drift} \sim \Big|\frac{r}{\dot{r}}\Big|,
\end{equation}
where $\dot{r}$ is the radial drift velocity given by Equation (\ref{eq:rdotpas}) for an irradiated disk and by Equation (\ref{eq:rdotact}) for an evolving disk.

\textit{Gas accretion timescale.} The timescale for gas accretion onto the central star for an evolving disk is (e.g., \citealt{armitage10})
\begin{equation}
\label{eq:tgas}
t_{\rm gas, acc} \sim \frac{r^2}{\nu} \sim \frac{1}{2 \alpha \eta \Omega_{\rm k}},
\end{equation}
with the latter expression derived from Equations (\ref{eq:nu}) and (\ref{eq:eta}).

@@

For simplicity, we calculate the radial drift timescale, $t_{\rm drift}$, for an irradiated disk in this section, but most of our conclusions hold true for an evolving disk as well. Figure \ref{fig:timescales} shows $t_{\rm des}$, $t_{\rm drift}$ and $t_{\rm gas, acc}$ as a function of particle size at three different locations in the disk, corresponding to the H$_2$O, CO$_2$ and CO snowlines in the static disk. As expected, micron-sized particles desorb on very short timescales of $\sim 1-1000$ years in the close vicinity of their respective snowlines, since the desorption rate depends exponentially on temperature and hence on disk location (see Equation \ref{eq:Rdes}).  On the other hand, their radial drift timescale exceeds the typical disk lifetime of a few Myr by several orders of magnitude due to their strong coupling with the gas. Thus for small particles in an irradiated disk, the snowline locations and the C/O ratio are the same as for a static disk (see Figure 1 from \citealt{oberg11}). This is not true for an evolving disk, however, where gas accretion causes even micron-sized particles to drift significantly before desorbing, as we show in section \ref{sec:snowlines}. At the other extreme, kilometer-sized particles are unaffected by gas drag and have long desorption timescales ($\gg$1 Myr ), and the snowline locations and C/O ratio remain unchanged in this case as well. This is true for both irradiated and evolving disks, since large planetesimals are decoupled from the gas and hence unaffected by gas accretion onto the host star.
